Uneven Investments in Toronto's Wards

### 2024-09-27 paper/paper.qmd
updated low income prevalence graph

# Create the table with kable and apply enhanced styling
kable(
  merged_table_2, 
  escape = TRUE,                   
  col.names = c(
    "Ward Name - Number", 
    "Per Capita Funding", 
    "Household Median Income", 
    "Low Income Prevalence", 
    "Bachelor’s Degree or Higher", 
    "Visible Minority Population"
  )
) |>
  kable_styling(
    bootstrap_options = c("condensed"),  
    full_width = FALSE,  
    position = "left",  
    font_size = 7 
  ) |>
  footnote(general =  "Rows in red indicate a ward where the Household Median Income is $84,000 or less.") |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "2cm") |>
  column_spec(3, width = "3cm") |>
  column_spec(4, width = "2cm") |>
  column_spec(5:6, width = "3cm") |>
  row_spec(1, color = "red") |>
  row_spec(6, color = "red") |>
  row_spec(9:10, color = "red") |>
  row_spec(13:14, color = "red") |>
  row_spec(17:19, color = "red") |>
  row_spec(21:24, color = "red")
```

Rows in **red** indicate a **Household Median Income** of \$84,000 or less.

@@

# Wards by Per Capita Funding
ggplot(merged_data, aes(x = reorder(`Ward`, per_capita_funding), y = per_capita_funding)) +
  geom_bar(stat = "identity", fill = "#0072B2") +
  labs(
    title = "Wards by Per Capita Funding",
    x = "Ward Name - Number",
    y = "Per Capita Funding ($CAD)"
  ) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal(base_family = "Helvetica") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title = element_text(size = 14),
    axis.text.y = element_text(size = 10)
  ) +
  coord_flip()
```

@@

# Combine these wards 
selected_wards <- bind_rows(top5_wards, bottom5_wards) |>
  mutate(Category = ifelse(`Ward Number` %in% top5_wards$`Ward Number`, "Top 5", "Bottom 5"))

@@

# Create the Combined Bar and Line Chart
ggplot(selected_wards, aes(x = reorder(`Ward Number`, per_capita_funding))) +
  geom_bar(aes(y = per_capita_funding, fill = Category), stat = "identity") +
  geom_line(aes(y = Scaled_Low_Income_Prevalence, group = 1, color = "Low Income Prevalence"), size = 1.2) +
  geom_point(aes(y = Scaled_Low_Income_Prevalence, color = "Low Income Prevalence"), size = 3) +

   scale_y_continuous(
    name = "Per Capita Funding ($CAD)",
    labels = dollar_format(),
    sec.axis = sec_axis(~./scale_factor, name = "Low Income Prevalence")
  ) +
  labs(
    title = "Top 5 and Bottom 5 Wards by Per Capita Funding with Low Income Prevalence",
    x = "Ward Number",
    fill = "Category",
    color = ""
  ) +

@@

  theme_minimal(base_family = "Helvetica") +
  theme(
    plot.title = element_text(size = 10, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.y.left = element_text(color = "black", size = 10),
    axis.title.y.right = element_text(color = "black", size = 10),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 8)
  )
```

>>>>

# Create the table with kable and apply enhanced styling
kable(
  merged_table_2, 
  escape = TRUE,                   
  col.names = c(
    "Ward Name - Number", 
    "Per Capita Funding", 
    "Household Median Income", 
    "Low Income Prevalence", 
    "Bachelor’s Degree or Higher", 
    "Visible Minority Population"
  )
) |>
  kable_styling(
    bootstrap_options = c("condensed"),  
    full_width = FALSE,  
    position = "left",  
    font_size = 7 
  ) |>
  footnote(general =  "Rows in red indicate wards with a Household Median Income at or below the average of $84,000.") |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "2cm") |>
  column_spec(3, width = "3cm") |>
  column_spec(4, width = "2cm") |>
  column_spec(5:6, width = "3cm") |>
  row_spec(1, color = "red") |>
  row_spec(6, color = "red") |>
  row_spec(9:10, color = "red") |>
  row_spec(13:14, color = "red") |>
  row_spec(17:19, color = "red") |>
  row_spec(21:24, color = "red")
```

@@

# Wards by Per Capita Funding
ggplot(merged_data, aes(x = reorder(`Ward`, per_capita_funding), y = per_capita_funding)) +
  geom_bar(stat = "identity", fill = "#0072B2") +
  labs(
    title = "Wards by Per Capita Funding",
    x = "Ward Name - Number",
    y = "Per Capita Funding ($CAD)"
  ) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title = element_text(size = 14),
    axis.text.y = element_text(size = 10)
  ) +
  coord_flip()
```

@@

# Combine these wards 
selected_wards <- bind_rows(top5_wards, bottom5_wards) |>
  mutate(Category = ifelse(`Ward` %in% top5_wards$`Ward`, "Top 5", "Bottom 5"))

@@

# Create the Combined Bar and Line Chart
ggplot(selected_wards, aes(x = reorder(`Ward`, per_capita_funding))) +
  geom_bar(aes(y = per_capita_funding, fill = Category), stat = "identity") +
  geom_line(aes(y = Scaled_Low_Income_Prevalence, group = 1, color = "Low Income Prevalence"), size = 1.2) +
  geom_point(aes(y = Scaled_Low_Income_Prevalence, color = "Low Income Prevalence"), size = 3) +

   scale_y_continuous(
    name = "Per Capita Funding ($CAD)",
    labels = dollar_format(),
    sec.axis = sec_axis(~./scale_factor, name = "Low Income Prevalence (Decimal)")
  ) +
  labs(
    title = "Top 5 and Bottom 5 Wards by Per Capita Funding with Low Income Prevalence",
    x = "Ward Name - Number",
    fill = "Category",
    color = ""
  ) +

@@

  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.y.left = element_text(color = "black", size = 10),
    axis.title.y.right = element_text(color = "black", size = 10),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 8)
  )
```
